% !TeX spellcheck = en_GB

\begin{comment}
\begin{itemize}
	\item experiment
	\item test data captured from a floor section of the computer science building
	\item enriched with malicious packets to keep consistent
	
	\item (focusses only a purpose based) attack classes (cf. \parencite{Uma2013})
		\subitem \gls{dos}
			\subsubitem Short circuit -> blackout on entire line
			\subsubitem flooding of \code{A\_Restart} telegrams
			\subsubitem flooding nonsense
		\subitem replay
			\subsubitem repeating a time window
			\subsubitem sniff a tag and repeat it compressed??? \alert{whatever this means?}
			\subsubitem do inverse action
		\subitem manipulation/reconfiguration
			\subsubitem telegram manipulation
			\subsubitem reconfiguration of devices (Access Attack)
			\subsubitem reconfigure line couplers/make them useless (Access Attack)
		\subitem spoofing 
		\subitem Reconnaissance Attack
			\subsubitem network mapping
			
		\subitem 
	
	\item aim is to show if attacks can be identified by anomaly detection on flow data
		\subitem under the assumption, that attacks noticeable alter the characteristic and behaviour of \gls{knx} traffic
		\subitem cf. \parencite{Mukherjee1994,Yang2006,Pan2014}
	\item demonstrate a message-passing architecture to perform online analytics on \gls{knx} flow-data
	\item benchmark different algorithms against each other
\end{itemize}
\end{comment}

\glsreset{ids}
The objective of this thesis is to investigate, whether anomaly based \glspl{ids} can provide additional security benefits for \gls{bas} in the same way they do for \gls{scada} networks. % (cf. Section~\ref{sec:intro})
Given the assumption that every malicious activity or attack within a network induces noticeable different characteristics and behaviour, these deviations might also be identifiable in aggregated flow-data of this network. \parencite[cf.][]{Mukherjee1994,Yang2006,Pan2014}
To investigate this hypothesis, a message-passing architecture to test different established algorithms for anomaly detection using flow-data from \gls{bas} networks was designed.

The investigation will be conducted as an experiment with data captured in a floor segment of the computer science building over the course of one month from 2017-01-21 to 2017-02-21. It is assumed, that this dataset does not contain any malicious activities.
This test dataset is divided into three parts:
The first two weeks of the dataset will be used to train multiple models using different algorithms.
Then the following third week is used as validation dataset to ensure that the models are not over-fitted and raise as few false-positives as possible.
Finally, the last week of the dataset is used as test dataset and is modified to include malicious activities, which might resemble a range of plausible attacks scenarios.
This, in turn, is done to ensure that the models and algorithms are not under-fitted and are suitable to detect anomalies, which might be induced by malicious activities.
However, the experiment will not focus on determining if the system can distinguish between malicious abnormal activities and legit abnormal activities, which might be caused by a rapid change of user behaviour. In this thesis both are considered worth reporting.

Mostly individual, small changes in user behaviour caused by an intruder are effectively indistinguishable from either actual completely normal behaviour or abnormal but allowed behaviour in the  here used test case of an University. Individual and small changes in the behaviour include for instance somebody turning on the lights at midnight. This cloud be a group of learning students, but it could also be an attacker installing malicious devices.
Due to this controversy the actual test and benchmarking will be done using a selection of plausible attacks an intruder might induce to gain control over the network.
%Mostly since individual, small changes in user behaviour are difficult to simulate in the here used test case, the actual test and benchmarking will be done using a selection of plausible attacks an intruder might induce.
Since real observations of attacks are not reported in literature and penetration testing tools targeting \gls{knx} or \gls{bas} in general are only sparsely available, the malicious traffic has to be simulated. In order to simulate the attacks artificially created packets are injected into the traffic datasets, which is described in more detail in Section~\ref{sec:methods:gen-test}

Plausible attacks include \emph{\glsfirst{dos} attacks}, which can be induced by multiple actions: First of all shorting the line circuit would effectively causing a communication blackout on the entire line. Also flooding \code{A\_Restart} telegrams to all devices would render them useless, since they would be stuck in a restart loop. As a last example of \gls{dos} attacks could be flooding the line with non-sense telegrams on high priority. (cf. Section~\ref{sec:background:bas:knx:proto})
This would cause at least heavily delayed, if not entirely blocked, normal communication on the line since the high priority telegrams would overrule all telegrams of lower priority.

Another plausible attack scenario would be \emph{replay attacks}, where three flavours might be imaginable.
In the first case an attacker would captures traffic for a determined time span and then replays it on the network, possibly to mimic normal behaviour while physically breaking into the building.
Secondly, an attacker could capture a specific event and replay it on will, which could contain the command to open a security door for instance.
As a third option, the traffic could be monitored and for every action a reverse action could be invoked, effectively keeping the \gls{bas} in one state. The simplest example would be, to turn off the lights every time somebody tries to turn them on.

The next category of attacks can be classified as manipulation or \emph{reconfigure attacks}. This includes attacks which might modify telegrams while they are send, to either render them invalid and therefore preventing communication. Also it would be conceivable to reconfigure devices, i.e. to trigger different actions, report false measurements etc. A specialisation of this attack focusses on line couplers and breaking the network segmentation by disabling any routing functions.

Finally, the last category describes \emph{reconnaissance attacks}. These attacks conclude unauthorised detection and mapping of the network and its behaviour. Here only active sweeping approaches are considered, where an attacker probes each individual device in an address range.
Passive eavesdropping is not considered as it can not be detected on higher protocol levels due to the bus character of the network. (cf. Section~\ref{sec:background:bas:knx:topo})

\section{Generating a Test Dataset including malicious activities}
\label{sec:methods:gen-test}

As \gls{bas} automation systems are only seldom considered within threat models, monitoring systems are rarely installed, if at all.
Consequently no sufficient amount of empirical data regarding attacks against \gls{bas} exists, simply because they would not be recognised as such, or simply because no one would publicly announce their vulnerability.
All of this leads to missing knowledge of what kind of attacks an intruder would perform in these networks.
Additionally, there are only a few tools available for performing penetration tests -- all of them are more targeted towards troubleshooting or reverse engineering the network topology.

However, to test the performance of the anomaly detection algorithms with regards to identifying malicious activities, data resembling those malicious activities or attacks is required.
Therefore, packets which imitate these activities are artificially crafted, then injected into the traffic, and later imported using a simulated Agent. (cf. Section~\ref{sec:impl:agent})
These activities follow the idea of the attack classes introduced above and are injected in the last week of the of the dataset.

First, unusual behaviour is injected at 2017-02-12 from 03:00 to 08:00, replacing this entire segment with traffic from 2017-02-06 10:00 to 15:00. In absence of data more foreign to the observed line the data was taken from the same line, but at a different weekday and day time. Effectively this data does not contain any new devices or severe unusual traffic, this modification rather aims to test the time sensitivity.

Second, a \gls{dos} attack is performed starting at 2017-02-13 09:00 and targeting the entire line \code{3.4}~. The attack is performed in three bursts of 15 minutes with five minutes break in between. In the \gls{dos} attack a flood of \code{A\_Restart} telegrams with \code{SYSTEM} priority is send, which in reality would cause all targeted devices to restart continuously. Additionally this blocks all other traffic, since the \code{SYSTEM} priority is the highest specified. During the attack the telegrams were injected with a maximum of \(500 \ \sfrac{telegrams}{min}\).

As third attack scenario a device scan over the entire possible \gls{knx} address space was performed, starting from 2017-02-13 21:00.
To determine if a device is present, the management \gls{apci}\break\code{A\_DEVICE\_DESCRIPTOR\_READ} is send to all addresses. Every \gls{knx} device is required to implement certain management routines, among them the query for the device descriptor. \code{A\_DEVICE\_DESCRIPTOR\_READ} is ideal since the requesting telegram does not require any parameters and the response only contains two bytes of additional payload. By choosing a request which is adds as little overhead as possible, the throughput is increased and effectively reducing the time required for the scan. \parencite[cf.][p.~46]{DIN_EN_50090-4-1}
Equal to the \gls{dos} attack, the telegrams are injected with a maximum of \(500 \ \sfrac{telegrams}{min}\).

Finally, two new rogue device are introduced with the addresses \code{3.6.26} and \code{3.5.18} during the entire day of 2017-02-14.
Again in lack of data which is foreign to this line but still resembles the activities of an ordinary \gls{knx} device, the traffic for the first device was taken from the address \code{3.6.7} at 2017-02-07 and for the second device from \code{2.6.42} at 2017-02-08.
The traffic was injected additional to the normal base-line activity on the target day, without any modification except for the source address and the change of the date.

The scripts to generate the malicious traffic and the datasets itself can be found on the data disk in Appendix~\ref{app:disk}.

\section{Evaluating the Detection Results}
\label{sec:methods:eval}

For each crafted attack, the different anomaly detection algorithms are benchmarked with regards to their ability to detect those.
This ability is classified by following criteria:

\begin{enumerate}
	\item General ability to detect the attack
	\item Differentiation from background noise of the detection results
		\begin{enumerate}
			\item Average difference in detected outliers
			\item Average difference in underlying decision metric
		\end{enumerate}
	\item Response time 
	\item Persistence of detection
\end{enumerate}

The first criteria describes, if the attack was detected at all by the respective anomaly detection algorithm.
However, the second criteria is meant to determine how well or clear the algorithm might distinguish between inlier or outlier. For this both the actual binary inlier or outlier decision as well as the underlying metric are considered. The only exception is the Entropy Analyser, since no threshold is predefined.
Third, the response time is measured from beginning of the attack to when a significant change in the measurements is shown.
Finally, the persistence of the detection is rated by determining the percentage of the time of the attack in which the activities are identified as outlier by the algorithm.

\begin{comment}
Angriffe:

DoS
	Kurzschluss im Bus -> DoS auf gesamtem Segment
	A_Restart-Pakete -> DoS gegen einzelne Teilnehmer
Replay-Angriffe
	Zeit mitschneiden -> wiedergeben
	Tag mitschneiden, komprimiert wiedergeben
Manipulation von Paketen (Payload tauschen)
Konfiguration manipulieren
Überwindung von Linienkopplern
Address-Spoofing
	falsche Adresse in Liniensegment
	mit existierender Adresse senden
Netzanalyse mit knxMap (https://github.com/takeshixx/knxmap)
Mitlesen und sofort gegenteilige Aktion auslösen
High-Level-Angriffe:
	nur best. Aktionen zulassen
	Provokation/Sabotage von menschl. Verhalten
Social-Engineering -> Einschleusen von Geräten
\end{comment}